<<<<<<< HEAD
This is markdown file for Peer Assessment 1 in Reproducible research course 
===========================================================================

## Loading and preprocessing the data

Read the activity.csv data into R object. We assume you download and unzip the data already and set working directory as well. 

```{r Read_file, echo=TRUE}
dataset <- read.csv("activity.csv")
```

Check the data.

```{r Check_data, echo=TRUE}
str(dataset)
sum(is.na(dataset[,1]))
```

## What is mean and median total number of steps taken per day?

##### Make a histogram of the total number of steps taken each day #####

Calculate total steps per day.
We use reshape2 and plyr packages for this.
We store the result into total_steps data frame. 

```{r library, echo=TRUE, results='hide'}
library(reshape2); library(dplyr)
```

```{r Steps/day, echo=TRUE}
total_steps_NA <- aggregate( steps ~ date, data=dataset, FUN=sum)
```

Create histogram of the total number of steps taken per day

```{r Steps/day hist, echo=TRUE}
rg_NA <- range(total_steps_NA$steps)
with(total_steps_NA, hist(steps, main="Histogram of total steps by date",
                            xlab = "Total steps by day",ylab = "Counts", col = "blue",
                            breaks=20, axes=FALSE))
axis(1, at=1000*0:rg_NA[2])
axis(2, at=2*0:20)

```

Calculate the mean and the median of total number of steps per day

```{r Steps/day mean, echo=TRUE}
M_steps_na <- aggregate( steps ~ date, data=dataset, FUN=mean)
Md_steps_na <- aggregate( steps ~ date, data=dataset, FUN=median)

steps_NA <- cbind(M_steps_na, Md_steps_na$steps)
names(steps_NA) <- c('Date', 'Mean of total steps/day', 'Median of total steps/day')
steps_NA
```

As you can see the values of medians are confusing. I tried many ways to calculate medians for each date according to the asingment but I didn't fixed it. 

## What is the average daily activity pattern?

Calculate avarage steps of intervals accross all days.
Remove the days without activity. 
We use reshape2 and plyr packages for this.
We store the result into steps_interval data frame. 

```{r Average steps, echo=TRUE}
steps_interval_mean <- aggregate( steps ~ interval, data=dataset, FUN=mean)
```

Make a plot of the average steps within interval 

```{r Average steps plot, echo=TRUE}
rg_I  <- range(steps_interval_mean$interval)
rg_I2 <- range(steps_interval_mean$steps)

with(steps_interval_mean, plot(interval, steps, type="n", lwd = 2, axes=FALSE, xlab="Interval", ylab="Average steps"))

axis(1, at=50*0:rg_I[2]) 
axis(2, at=20*0:rg_I2[2])
with(steps_interval_mean, lines(interval, steps, col="red"))
title(main='Activity measure')
```

*Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?* 

```{r Max steps interval, echo=TRUE}
steps_interval_max <-  aggregate(steps ~ interval, data=dataset, FUN=sum)

steps_interval_max <- steps_interval_max[order(steps_interval_max[2], decreasing=T),]

max_i <- steps_interval_max[1,]; names(max_i) <- c('ID of interval', 'Number of steps'); print(max_i)
```

The 5-minute interval which contains the maximum number of steps is interval number  `r max_i[[1]]`. 
=======
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?

>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270


## Imputing missing values

<<<<<<< HEAD
Calculate the total number of missing values in the dataset 

```{r Total NAs, echo=TRUE}
sum(is.na(dataset))
```

Find NA value in dataset and replace them by average number of steps within the interval accross all days.

```{r NAs replacement, echo=TRUE}

interval <- unique(dataset$interval)
for (i in interval) {
      dataset[dataset$interval==i & is.na(dataset$steps), "steps"] <- 
      steps_interval_mean[steps_interval_mean$interval==i, "steps"]      
}

sum(is.na(dataset)) ## check 
```

Calculate the total sum of steps each day

```{r Steps/day without NAs, echo=TRUE}
total_steps <- aggregate(steps ~ date, data=dataset, FUN=sum)
```

Create histogram of the total number of steps taken per day.

```{r Steps/day hist without NAs, echo=TRUE}
rg <- range(total_steps$steps)
with(total_steps, hist(steps, main="Histogram of total steps by date",
                            xlab = "Total steps by day",ylab = "Counts", col = "red",
                            breaks=20, axes=F))
axis(1, at=1000*0:rg[2])
axis(2, at=2*0:20)
```

Calculate mean and median of steps taken per day

```{r M/Md , echo=TRUE}
M_steps <- aggregate(steps ~ date, data=dataset, FUN=mean)

Md_steps <- aggregate(steps ~ date, data=dataset, FUN=median)

MMd_steps <- cbind(M_steps, Md_steps$steps)

names(MMd_steps) <- c('Date', 'Mean of total steps/day', 'Median of total steps/day')
MMd_steps
```

Now we can compare the results between the first and third part of assingment. 

```{r Compare Means and Medians , echo=TRUE}
## compare means ## compare means ## compare means 

compare_mean <- merge(M_steps_na, M_steps, by="date", all.y=T)
names(compare_mean) <- c('Date', 'Without NA correction', 'With NA correction')
compare_mean

## compare medians ## compare medians ## compare medians 

compare_median <- merge(Md_steps_na, Md_steps, by="date", all.y=T)
names(compare_median) <- c('Without NA correction', 'With NA correction')
compare_median
```


## Are there differences in activity patterns between weekdays and weekends?

We use weekdays() fuction and filter() function from dplyr package to subset and compare activity from weekdays and weekends. 

* Please note that names of weekdays are in Slovak language 

```{r Compare weekdays and weekends , echo=TRUE}
dataset$date <- as.Date(as.character(dataset$date)) ## set date variable to Date class

## subset weekdays activity 
df_weekdays <- filter(dataset, weekdays(date)=='pondelok'| weekdays(date)=='utorok' | weekdays(date)=='streda'| weekdays(date)=='štvrtok'| weekdays(date)=='piatok')

## subset weekend activity 
df_weekend <- filter(dataset, weekdays(date)=='sobota'| weekdays(date)=='nede¾a')

weekdays_avg <- aggregate(steps ~ interval, data=df_weekdays, FUN=mean)
 
weekends_avg <- aggregate(steps ~ interval, data=df_weekend, FUN=mean)
```

```{r Plot weekdays/weekends, echo=TRUE}
par(mfrow = c(2,1), mar=c(5,4,1,1), las=1)

## Plot for Weekdays 
rg_d1  <- range(weekdays_avg$interval)
rg_d2 <- range(weekdays_avg$steps)

with(weekdays_avg, plot(interval, steps, type="n", axes=F, xlab="Interval", ylab="Average steps"))
axis(1, at=150*0:rg_d1[2]) 
axis(2, at=30*0:rg_d2[2])
with(weekdays_avg, lines(interval, steps, col="chartreuse4", lwd = 2))
title(main='Actitivity measure during weekdays')

## Plot for Weekends
rg_e1  <- range(weekends_avg$interval)
rg_e2  <- range(weekends_avg$steps)

with(weekends_avg, plot(interval, steps, type="n", axes=F, xlab="Interval", ylab="Average steps"))
axis(1, at=150*0:rg_d1[2]) 
axis(2, at=30*0:rg_d2[2])
with(weekends_avg, lines(interval, steps, col="cadetblue4", lwd = 2))
title(main='Actitivity measure during weekends')

```
=======


## Are there differences in activity patterns between weekdays and weekends?
>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270
